|

### 2021-02-23 appendix/appendix.Rmd
small edits to appendix and manuscript

Workers perform preventive healthcare tasks. To inspect a house and eliminate mosquito breeding sites, healthcare workers visit the households in pairs or in groups of three. Supposedly, this increases efficiency as it helps the team cover more territory and be more careful in the areas they inspect individually. Team structures also help in terms of lowering free-riding and increasing incentives for cooperative work. To capture this, we assume that a given worker $i$ divides her work effort into two tasks: those which she accomplishes with her own effort ($e_i$) and helping efforts ($h_i$). To capture the costs of effort, we assume the following convex cost function $\dfrac{c}{2}(e_i^2 + h_i^2)$, where $c>0$ represents the changes in marginal costs of effort and help.

@@

## Baseline: No Incentive Bonuses

@@

The prediction here is that workers will exert no effort, and that the agency will have only costs without accruing any benefit. This is the worst-case scenario in terms of the agency's welfare. In the experiment, it matches the control condition.\footnote{Throughout the chapter, we assume that the optimal output for the agency is equal to the maximum welfare for society. Although this is not always the case, it is a reasonable simplification for the purposes of this article. Adding a difference here would only complicate our results and provide no further insight into designing an incentive scheme to maximise welfare.}

@@

We have five data sources in this study, all of them are included in `./data/incentives_healthcare.RData`:

- `pretreat`: Pre-treatment balance information.
- `prodfr`: Field productivity data.
- `dfr`: Disease incidence in every Rio Verde census sector.
- `dfr2`: Disease incidence in every Rio Verde census sector that received the intervention.
- `did_full`: Differences-in-differences estimator that measures the effect of our intervention against a pure control group.

@@

The pre-treatment balance dataset contains information about every individuals who interacted with the experimental team at any point of time. To be included in this dataset, the person has to have filled the recruitment form we had circulated on Facebook as part of our campaign. The dataset contains 13 variables that we collected before the intervention. We gathered these data either via Google Forms, or in pre-treatment surveys before the training session.

@@

We collected field productivity data based on the subjects' performance during the experiment. We measured productivity using three indicators:

@@

We compiled these data using pictures taken in the field. All participants were instructed to take pictures of clean breeding sites, and to make videos showing the larvae they exterminated. After finishing the inspection of each household, subjects had to take a last picture of the house to prove that they had been there.

@@

1. `D`: Treatment assignment.
2. `pair_n`: Pair index ID.
3. `nHouse`: Number of houses visited.
4. `nBreed`: Number of breeding sites cleaned.
5. `nLarvae`: Number of larvae exterminated.
6. `nHouseVisit`: Number of houses that were indeed visited (based on having pictures of breeding sites and larvae videos).
7. `nHouseLessTwoMin`: Number of houses visited in less than two minutes (which denotes either that the pair did not go to the place, or that there were no residents at home).

@@

To assess whether the intervention had an impact on the municipal healthcare system, we collected thirteen variables in two distinct datasets. All datasets compile data from the National Health Notification System (SINAN) for the  Rio Verde municipality. The dataset has the complete records of all individuals who had been hospitalised between four months before and four month after the intervention. We geolocated the household addresses of those individuals who were hospitalised with dengue fever, then we matched the case to the census sector where that person lives. Finally, we computed the number of cases before and after a few time windows around the intervention. The result is two datasets, `dfr` (with all census sectors) and `dfr2` (with only the census sectors matched in the intervention). Both datasets have the following variables:

1. `CensusSectors`: Last three digits of census sector (2010 Brazilian census).
2. `NHouses`: Number of houses in each census sector (2010 Brazilian census).
3. `NHouseholds`: Average Number of households in each house/building and in each census sector (2010 Brazilian Census).
4. `AvgHouseholdSize`: Average household income per census sector (2010 Brazilian Census).
5. `logAvgIncomeHousehold`: Log of average household income per census sector (2010 Brazilian Census).
6. `D1`: Treatment assignment.
7. `D2`: Intervention x no intervention (no Intervention empty in `dfr2`).
8. `nc15dafter`: Number of cases 15 days after the intervention.
9. `nc30dafter`: Number of cases 30 days after the intervention.
10. `nc60dafter`: Number of cases 60 days after the intervention.
11. `nc90dafter`: Number of cases 90 days after the intervention.
12. `nc30to60dafter`: Number of cases between 30 and 60 days after the intervention.
13. `plac`: Number of cases before the intervention.

@@

We build `did_full` dataset to investigate whether the intervention had any impact on public health outcomes. The disease incidence data describes the treatment effect. In the differences-in-differences models, we combine all treatments and investigate how they perform against a pure control group, which are the census sectors that were not visited during the intervention.

@@

We measure covariate balance using four statistics. First, we check whether subjects decided to join us for the training day. Second, we assess their decision to participate in the experiment. Subjects had to take part in the training process and then participate in the fieldwork. Third, we checked if individual characteristics are balanced across our treatment groups. We collected that information via Google Forms and using a pre-treatment survey. Finally, we checked the territorial assignment to make sure that the pre-treatment characteristics are balanced in the places that subjects visited. We find a good level of balance for the individual assignment (models 1 to 3), and we corrected eventual imbalances in the territorial assignment using propensity-score matching [@rosenbaum1985constructing].

@@

## How and When Manipulations were Administered

@@

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth, height=10cm]{../figures/rvFieldPic2.jpg}
\caption{Field pictures}
\end{figure}

@@

# R Session Information

>>>>

Workers perform preventive healthcare tasks. To inspect a house and eliminate mosquito breeding sites, healthcare workers visit the households in pairs or in groups of three. Supposedly, this increases efficiency as it helps the team cover more territory and be more careful in the areas they inspect individually. Team structures also help in terms of lowering free-riding and increasing incentives for cooperative work. To capture this, we assume that a given worker $i$ divides her work effort into two tasks: her own effort ($e_i$) and helping efforts ($h_i$). To capture the costs of effort, we assume the following convex cost function $\dfrac{c}{2}(e_i^2 + h_i^2)$, where $c>0$ represents the changes in marginal costs of effort and help.

@@

## Baseline: No Bonus

@@

The prediction here is that workers will exert no effort, and that the agency will have only costs without accruing any benefit. This is the worst-case scenario in terms of the agency's welfare. In the experiment, it matches the control condition.\footnote{Throughout the article, we assume that the optimal output for the agency is equal to the maximum welfare for society. Although this is not always the case, it is a reasonable simplification for the purposes of this article. Adding a difference here would only complicate our results and provide no further insight into designing an incentive scheme to maximise welfare.}

@@

We use five data sources in this study. All of them are included in `../data/incentives_healthcare.RData`:

- `pretreat`: Pre-treatment balance information.
- `prodfr`: Field productivity data.
- `dfr`: Disease incidence in every Rio Verde census sector.
- `dfr2`: Disease incidence in every Rio Verde census sector that received the intervention.
- `did_full`: Differences-in-differences estimation that measures the effect of our intervention against a pure control group.

@@

The pre-treatment balance dataset contains information about every individual who interacted with the research team at any point of time. To be included in this dataset, the person needs to have filled the Google Form we circulated on Facebook as part of our recruitment campaign. This dataset contains 13 variables that we collected before the intervention. We gathered these data either via Google Forms, or in pre-treatment surveys before the training session.

@@

We collected field productivity data based on the subjects' performance during the experiment. We measured productivity with three indicators:

@@

We compiled these data using photos taken during the field experiment. All participants were instructed to take pictures of clean breeding sites, and to make videos showing the larvae they exterminated. After finishing the inspection of each household, subjects had to take a last picture of the house to prove that they had been there.

@@

1. `D`: Treatment assignment.
2. `pair_n`: Pair index ID.
3. `nHouse`: Number of houses visited.
4. `nBreed`: Number of breeding sites cleaned.
5. `nLarvae`: Number of larvae exterminated.
6. `nHouseVisit`: Number of houses that were indeed visited (based on pictures of breeding sites and larvae videos).
7. `nHouseLessTwoMin`: Number of houses visited in less than two minutes (which denotes either that the pair did not go to the place, or that there were no residents at home).

@@

To assess whether the intervention had an impact on the municipal healthcare system, we collected 13 variables in two distinct datasets. All datasets compile data from the National Health Notification System (SINAN) for the Rio Verde municipality. The dataset has the complete records of all individuals who had been hospitalised between four months before and four month after the intervention. We geolocated the household addresses of those individuals who were hospitalised with dengue fever, then we matched the case to the census sector where that person lives. Finally, we computed the number of cases before and after a few time windows around the intervention. The result is two datasets, `dfr` (with all census sectors) and `dfr2` (with only the census sectors matched in the intervention). Both datasets have the following variables:

1. `CensusSectors`: Last three digits of census sector (2010 Brazilian census).
2. `NHouses`: Number of houses in each census sector (2010 Brazilian census).
3. `NHouseholds`: Average Number of households in each house/building and in each census sector (2010 Brazilian Census).
4. `AvgHouseholdSize`: Average household income per census sector (2010 Brazilian Census).
5. `logAvgIncomeHousehold`: Log of average household income per census sector (2010 Brazilian Census).
6. `D1`: Treatment assignment.
7. `D2`: Intervention x no intervention (no intervention missing in `dfr2`).
8. `nc15dafter`: Number of cases 15 days after the intervention.
9. `nc30dafter`: Number of cases 30 days after the intervention.
10. `nc60dafter`: Number of cases 60 days after the intervention.
11. `nc90dafter`: Number of cases 90 days after the intervention.
12. `nc30to60dafter`: Number of cases between 30 and 60 days after the intervention.
13. `plac`: Number of cases before the intervention.

@@

We build the `did_full` dataset to investigate whether our intervention had any impact on public health outcomes. The disease incidence data describe the treatment effect. In the differences-in-differences models, we combine all treatments and investigate how they perform against a pure control group, which are the census sectors that were not visited during the intervention.

@@

We measure covariate balance using four statistics. First, we check whether subjects decided to take part in the training sessions. Second, we evaluate the balance among those who decided to participate in the experiment. Subjects had to take part in the training process and participate in the fieldwork. Third, we check if individual characteristics are balanced across our treatment groups. As we mentioned above, we collected that information via Google Forms and using a pre-treatment survey. Finally, we check the territorial assignment to make sure that the pre-treatment characteristics are balanced across the places that subjects visited. We find good balance for the individual assignment (models 1 to 3), and we corrected eventual imbalances in the territorial assignment using propensity score matching [@rosenbaum1985constructing].

@@

## How and When Manipulations Were Administered

@@

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth, height=13cm]{../figures/rvFieldPic2.jpg}
\caption{Field pictures}
\end{figure}

@@

# Session Information

### 2021-02-23 manuscript/article.Rmd
small edits to appendix and manuscript

---
title:  |
        | Financial Incentives and Healthcare Provision: Evidence from an Experimental *Aedes aegypti* Control Programme in Brazil
author:
  - Danilo Freire[^freire]
  - Umberto Mignozzetti[^mignozzetti]
date: \today
thanks: "We thank Livio DiLonardo, Saad Gulzar, Cathy Hafer, Carlo Holz, Dimitri Landa, Guisela Pereira, Adam Przeworski, Catarina Roman, David Skarbek, Denis Stukal, Marcos Tourinho, Scott Tyson, and participants at MPSA, NYU PE Seminar, and SPSA for their valuable comments. Replication materials are available at <http://github.com/danilofreire/incentives-health-care>."
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: unsrtnat
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: template.latex
---

@@

\noindent \textbf{Methodology/Principal findings}. We hired and trained
subjects to visit households, find \textit{A. aegypti} breeding sites, and
eliminate mosquito larvae in the city of Rio Verde, Brazil. We randomly
assigned workers into three groups. The control group received a flat
compensation for their tasks, while workers in the two treatment groups received
individual and collective monetary bonuses, respectively. Financial rewards
increased the number of cleaned breeding sites in both treatment groups
(individual and team bonuses) and the collective treatment also improved larvae
extermination. The intervention lowered dengue hospitalisations in 10.3\%, but
the result was not consistent across all model specifications.

\noindent \textbf{Conclusions/Significance}. \textit{A. aegypti} control
programmes may benefit from alternative compensation schemes, specially when
provided to teams. For this strategy to succeed, financial incentives have to
be distributed widely as their aggregate effect is limited. More research is
needed to assess whether higher worker productivity decreases dengue
hospitalisations.

@@

\noindent Diseases transmitted by the \textit{Aedes aegypti} mosquito, such as
chikungunya, dengue, yellow fever, and Zika, continue to affect thousands of
people per year. As there are no safe vaccines for most of these infections,
insecticide spraying and breeding site elimination are the best means to fight
the mosquito. In several developing countries, which host the majority of
\textit{A. aegypti} infections, anti-mosquito campaigns are carried out
inconsistently, thus it is crucial to find ways to improve the productivity of
healthcare workers in charge of these tasks. We designed a randomised field
experiment that provided individual and collective financial incentives to
healthcare agents in a Brazilian city, and we tested the effect of monetary
rewards on their productivity and on city-level dengue hospitalisations. We
find that financial bonuses improved the number of cleaned breeding sites in
both treatment groups (individual and team incentives) and that the collective
treatment also improved larvae extermination. The impact of our treatment on
city-level hospitalisations was not consistent across all specifications. In
sum, financial incentives may be used to boost field productivity in
anti-\textit{A. aegypti} programmes, but further research is required to
evaluate how healthcare worker productivy impacts dengue outcomes.

@@

Mosquito control is the primary strategy to prevent *Aedes*-borne infections
[@beatty2010best; @eisen2009proactive]. The World Health Organization
recommends the cleaning of containers and the use of insecticide to reduce
vector breeding, specially in settings where human-vector contacts are
frequent, such as schools, hospitals, and workplaces [@WHO2012global;
@world2020denguecontrol]. However, breeding site elimination and insecticide
spraying are labour-intensive processes, which are often carried out
inconsistently by developing states [@paz2015dengue; @siddiqui2016use;
@WHO2012global]. In this regard, improving their health agency's performance in
mosquito control is essential to the effectiveness of anti-*Aedes* programmes.

@@

In this paper, we explore how financial incentives affect the productivity of
healthcare personnel involved in *A. aegypti* control and how they impact
dengue hospitalisations. We ran a randomised field experiment in the city of
Rio Verde, Brazil, which has experienced a peak in dengue infections in recent
years. We hired and trained teams of healthcare workers to perform simplified
procedures designed to exterminate *A. aegypti* breeding sites. Our experiment
consisted of two treatment arms plus a control condition. We offered individual
or collective bonuses to those allocated in the two treatment groups, while the
control group received a fixed compensation for completing their tasks. We then
measured how the different reward schemes affected the number of breeding sites
discovered, larvae exterminated, and city-level numbers of dengue incidence.

@@

Table \ref{tab:covbal} displays the attrition rates and the covariance balance
tests for our experiment. The last two tables report the values for our
territorial assignment before and after applying propensity score matching
[@rosenbaum1985constructing] on census tract characteristics. We observe no
significant differences regarding attrition rates, indicating that the
treatment was correctly randomised. Participant demographics are also similar
across all groups, what suggests that the matching technique successfully
balanced the territorial assignment.

>>>>

---
title:  |
        | Financial Incentives and Healthcare Provision: Evidence from an Experimental *Aedes aegypti* Control Programme in Brazil
author:
  - Danilo Freire[^freire]
  - Umberto Mignozzetti[^mignozzetti]
date: \today
thanks: "We thank Saad Gulzar, Cathy Hafer, Dimitri Landa, Rebecca Morton, Adam Przeworski, Catarina Roman, Renard Sexton, David Skarbek, Denis Stukal, and participants at EGAP, MPSA, NYU PE Seminar, and SPSA for their valuable comments. We also thank the FGV Applied Research Centre for funding this research. Replication materials are available at <http://github.com/danilofreire/incentives-healthcare>."
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: unsrtnat
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: template.latex
---

@@

\noindent \textbf{Methodology/Principal findings}. We hired and trained
subjects to visit households, find \textit{A. aegypti} breeding sites, and
eliminate mosquito larvae in the city of Rio Verde, Brazil. We randomly
assigned workers into three groups. The control group received a flat
compensation for their tasks, while workers in the two treatment groups received
individual and collective monetary bonuses, respectively. Financial rewards
increased the number of cleaned breeding sites in both treatment groups
(individual and team bonuses), and the collective treatment also improved larvae
extermination. The intervention lowered dengue hospitalisations in 10.3\%, but
the result was not consistent across all model specifications.

\noindent \textbf{Conclusions/Significance}. \textit{A. aegypti} control
programmes may benefit from alternative compensation schemes, especially when
provided to teams. For this strategy to succeed, financial incentives have to
be distributed widely as their aggregate effect is limited. More research is
needed to assess whether higher worker productivity decreases dengue
hospitalisations.

@@

\noindent Diseases transmitted by the \textit{Aedes aegypti} mosquito, such as
chikungunya, dengue, yellow fever, and Zika, continue to affect thousands of
people per year. As there are no safe vaccines for most of these infections,
insecticide spraying and breeding site elimination are the best means to fight
the mosquito. In several developing countries, which host the majority of
\textit{A. aegypti} infections, anti-mosquito campaigns are carried out
inconsistently, thus it is crucial to find ways to improve the productivity of
healthcare workers in charge of these tasks. We designed a randomised field
experiment that provided individual and collective financial incentives to
healthcare agents in a Brazilian city, and we tested the effect of monetary
rewards on their productivity and on city-level dengue hospitalisations. We
find that financial bonuses improved the number of cleaned breeding sites in
both treatment groups (individual and team incentives) and that the collective
treatment also improved larvae extermination. The impact of our treatment on
city-level hospitalisations was not consistent across all specifications. In
sum, financial incentives may be used to boost field productivity in
anti-\textit{A. aegypti} programmes, but further research is required to
evaluate how healthcare worker productivity impacts dengue outcomes.

@@

Mosquito control is the primary strategy to prevent *Aedes*-borne infections
[@beatty2010best; @eisen2009proactive]. The World Health Organization
recommends the cleaning of containers and the use of insecticide to reduce
vector breeding, especially in settings where human-vector contacts are
frequent, such as schools, hospitals, and workplaces [@WHO2012global;
@world2020denguecontrol]. However, breeding site elimination and insecticide
spraying are labour-intensive processes, which are often carried out
inconsistently by developing states [@paz2015dengue; @siddiqui2016use;
@WHO2012global]. In this regard, improving their health agency's performance in
mosquito control is essential to the effectiveness of anti-*Aedes* programmes.

@@

In this paper, we explore how financial incentives affect the productivity of
healthcare personnel involved in *A. aegypti* control and how they impact
dengue hospitalisations. We ran a randomised field experiment in the city of
Rio Verde, Brazil, which has experienced a peak in dengue infections in recent
years. We hired and trained teams of healthcare workers to perform simplified
procedures designed to exterminate *A. aegypti* breeding sites. Our experiment
consisted of two treatment arms plus a control condition. We offered individual
or collective bonuses to those allocated in the two treatment groups, while the
control group received a fixed compensation for completing their tasks. We then
measured how the different reward schemes affected the number of breeding sites
discovered, larvae exterminated, and city-wide dengue incidence.

@@

Table \ref{tab:covbal} displays the attrition rates and the covariance balance
tests for our experiment. The last two tables report the values for our
territorial assignment before and after applying propensity score matching
[@rosenbaum1985constructing] on census tract characteristics. We observe no
significant differences regarding attrition rates, indicating that the
treatment was correctly randomised. Participant demographics are also similar
across all groups, which suggests that the matching technique successfully
balanced the territorial assignment.
